---
title: "p8105_hw1_hx2264"
author: "Hongyao Xie"
date: "September 13, 2018"
output: github_document
---

#Problem 1 

```{r chunk1}

#load the library tidyverser
library(tidyverse)

#Set seed to get a reproducible random result
set.seed(1)

#create the data frame
df_4_var = tibble(
  sample = runif(10, min = 0, max = 5),
  sample_over_2 = sample > 2,
  char = as.character(rep(c("a","b","c","d","f"),2)),
  fac = as.factor(rep(c("a","b","c","d","f"),2))
) 

#run the data frame to see if it works
head(df_4_var)

#take the mean of each variable
mean(df_4_var$sample)
mean(df_4_var$sample_over_2)
mean(df_4_var$char)
mean(df_4_var$fac)
```

The column 1 and 2 (ramdom sample and logical vector) worked, whereas the column 3 and 4 (character vector and factor vector) didn't. Mean of the random sample was its arithmetic mean, and mean of the logical vector was to regard TRUE as 1 and FALSE as 0. Both columns could be interpreted as numeric values, and thus we could calculate their mean. 

The column 3 and 4 didn't work because character and factor could not be interpreted as numeric values. Any words, like "hello" or "aaa" or "123", could be used in character vector and factor vector. Since the elements in these vectors were not always numbers, these two variable types did't have mean.



```{r chunk2, results = hide}
#applies as.numeric function
as.numeric(df_4_var$sample_over_2)
as.numeric(df_4_var$char)
as.numeric(df_4_var$fac)
```

The as.numeric function applied to logical and factor variables but the character variable. For the logical variable, as.numeric function returned 0 for FALSE and returned 1 for TRUE. For the factor variable, as.numeric returned an integer representation for each level of the factor variable. However, for the character variable, as.numeric wouldn't return any numeric values.


```{r chunk3}
#convert character variable from character to factor to numeric
char_to_fac <- as.factor(df_4_var$char)
char_to_fac_to_num <- as.numeric(char_to_fac)
char_to_fac_to_num

#convert factor variable from factor to character to numeric
fac_to_char <- as.character(df_4_var$fac)
fac_to_char_to_num <- as.numeric(fac_to_char)
fac_to_char_to_num
```



#Problem 2

```{r}
#create the data frame:
#dimision of the data fram is 1000 rows and 5 columns
set.seed(2)

df_5_var = tibble(
  x = rnorm(1000),
  y = rnorm(1000),
  sum_over_0 = x + y > 0,
  sum_over_0_to_num = as.numeric(sum_over_0),
  sum_over_0_to_fac = as.factor(sum_over_0)
)

dim(df_5_var)
head(df_5_var)

#mean and median of x
mean(df_5_var$x)
median(df_5_var$x)

#the proportion of cases for which the logical vector is TRUE
length(which(df_5_var$sum_over_0))/length(df_5_var$sum_over_0)
```


```{r}
library(gridExtra)

#scatterplot of y vs x, color using the logical variable
scatter_by_logical <- ggplot(df_5_var, 
                             aes(x = x, y = y, color = sum_over_0)) + geom_point()

ggsave("scatter_by_logical.png")

#scatterplot of y vs x, color using the numeric variable
scatter_by_num <- ggplot(df_5_var, 
                         aes(x = x, y = y, color = sum_over_0_to_num)) + geom_point()

ggsave("scatter_by_num.png")

#scatterplot of y vs x, color using the factor variable
scatter_by_fac <- ggplot(df_5_var, 
                         aes(x = x, y = y, color = sum_over_0_to_fac)) + geom_point()

ggsave("scatter_by_fac.png")

#combine multiple plots
grid.arrange(scatter_by_logical, scatter_by_num, scatter_by_fac, ncol = 2)

ggsave
```




















